Generating gaits for physical quadruped robots: evolved neural networks vs. local parameterized search

### 2011-04-13 paper_ecal/experimentalSetup.tex
More edits, 75% done :)

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

@@

\input{platformDetails}

\input{fitnessEvaluation}

>>>>

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world instead of
using a simulator.  In this section we give details of the hardware
setup we used.  First we discuss the robotic platform on which gaits
were run, and then we describe the tracking system used to quantify
the fitness of those gaits.

%: \secref{platformDetails} discusses the robotic
%platform, and \secref{fitnessEvaluation} describes the tracking system
%we used to measure each gait's fitness.

@@

%\input{platformDetails}

%\input{fitnessEvaluation}

### 2011-04-13 paper_ecal/hyperNeatMethod.tex
More edits, 75% done :)

\acmFig{hyperneatExplanation.png}{.8}{HyperNEAT produces ANNs from
  CPPNs. ANN weights are specified as a function of the geometric
  coordinates of the source node and the target node for each
  connection. The coordinates of these nodes and a constant bias are
  iteratively passed to the CPPN to determine each connection
  weight. The CPPN has two output values, which specify the weights
  for each connection layer as shown.}

@@

\acmFig{SpiderANN.jpg}{.7}{ANN configuration for HyperNEAT Runs. The
  first two columns of each row of the input layer receive information
  about a single leg (the current angle of its two joints). The final
  column provides a sine and cosine wave to enable periodic movements
  and the angle of the center joint. Evolution determines the function
  of the hidden-layer nodes. The nodes in the output layer specify new
  joint angles for each respective joint. The unlabeled nodes in the
  input and output layers are ignored.}\label{hyperneatLayout}

>>>>

\acmFig{hyperneatExplanation.png}{.8}{HyperNEAT produces ANNs from
  CPPNs. ANN weights are specified as a function of the geometric
  coordinates of the source node and the target node for each
  connection. The coordinates of these nodes and a constant bias are
  iteratively passed to the CPPN to determine each connection
  weight. The CPPN has two output values, which specify the weights
  for each connection layer as shown.  Figure from
  \cite{clune2011performance}.}

@@

\acmFig{SpiderANN.jpg}{.7}{ANN configuration for HyperNEAT Runs. The
  first two columns of each row of the input layer receive information
  about a single leg (the current angle of its two joints). The final
  column provides a sine and cosine wave to enable periodic movements
  and the angle of the center joint. Evolution determines the function
  of the hidden-layer nodes. The nodes in the output layer specify new
  joint angles for each respective joint. The unlabeled nodes in the
  input and output layers are ignored. Figure adapted from
  \cite{clune2011performance}.}\label{hyperneatLayout}

### 2011-04-13 paper_ecal/learningMethods.tex
More edits, 75% done :)

@@

We have chosen to implement two classes of algorithms to automatically create the motion-generating function $g(t)$ (described in
\secref{problemDefinition}).  The first method uses a family of
parameterized gaits, described in this section and the next. The
second method evolves a non-parametric gait controlled by neural networks, and is detailed in
\secref{hyperNeatMethod}.

@@

Using the SineModel5 parameterized motion model described in the
previous section, along with the allowable ranges for each of the five
parameters (shown in \tabref{parameters}), the task becomes how to
choose the combination of five parameters that results in the fastest
motion, per the evaluation methods in \secref{fitnessEvaluation}.

>>>>

We have chosen to implement two classes of algorithms to produce
the motion-generating function $g(t)$ (described earlier).  The
first method uses a family of parameterized gaits, described in this
section and the next. The second method, HyperNEAT, evolves a non-parametric gait
controlled by neural networks and is described last.

@@

@@

Using the SineModel5 parameterized motion model described in the
previous section, along with the allowable ranges for each of the five
parameters (shown in \tabref{parameters}), the task becomes how to
choose the combination of five parameters that results in the fastest
motion, per the chosen evaluation methods described earlier.

### 2011-04-13 paper_ecal/main.tex
More edits, 75% done :)

% Section 3
\section{Experimental Setup}
\seclabel{experimentalSetup}

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

\input{platformDetails}
\input{fitnessEvaluation}

>>>>

% Section 3
\input{experimentalSetup}
\input{platformDetails}
\input{fitnessEvaluation}

### 2011-04-13 paper_ecal/problemDefinition.tex
Moved GECCO paper to its own directory, copied to new directory for ECAL paper
Reworded a few sections
More edits, 75% done :)

\section{Problem definition}
\seclabel{problemDefinition}

% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
%
%\edit{write this.  Quadratot report section below}
%
%We are testing several different learning methods to design a
%parametrized gait for a quadruped robot from the Cornell Computational
%Synthesis Lab.
%
%The output each of the learning algorithms is a function of time,
%$g(t)$, that outputs a vector of commanded motor positions.  This
%function is generated using a parametrized motion model, described in
%\secref{implement}.
%
%The robot executes these commands and measures its change in location
%using the tracking system described in \secref{implement}.  The input
%to the learning algorithms is this measured displacement, which the
%algorithms attempt to maximize. This displacement is measured for each
%gait over a constant length run, usually 12 seconds.
%
%A comparison and evaluation of the many different methods available
%for optimizing the gait of legged robots will be useful for future
%work on this challenging multidimensional control problem.

@@

For the system evaluated in this paper, we chose to use only
feedforward gaits.  An interesting extension would be to allow gaits
to depend on the measured servo positions, loads, voltage drops, or
other quantities.

The ultimate goal was to design gaits that were as fast as possible,
so to optimize for speed, our performance metric was displacement over
a period of motion lasting approximately 12 seconds.  For details of
exactly how this displacement was measured, see
\secref{fitnessEvaluation}.

>>>>

\section{Problem Definition}
\seclabel{problemDefinition}

% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
%
%\edit{write this.  Quadratot report section below}
%
%We are testing several different learning methods to design a
%parameterized gait for a quadruped robot from the Cornell Computational
%Synthesis Lab.
%
%The output each of the learning algorithms is a function of time,
%$g(t)$, that outputs a vector of commanded motor positions.  This
%function is generated using a parameterized motion model, described in
%\secref{implement}.
%
%The robot executes these commands and measures its change in location
%using the tracking system described in \secref{implement}.  The input
%to the learning algorithms is this measured displacement, which the
%algorithms attempt to maximize. This displacement is measured for each
%gait over a constant length run, usually 12 seconds.
%
%A comparison and evaluation of the many different methods available
%for optimizing the gait of legged robots will be useful for future
%work on this challenging multidimensional control problem.

@@

For the system evaluated in this paper, we chose to compare only
feedforward gaits generated by both the parameterized methods and
HyperNEAT.  An interesting extension would be to allow gaits to depend
on the measured servo positions, loads, voltage drops, or other
quantities.

The ultimate goal was to design gaits that were as fast as possible,
so to optimize for speed, our performance metric was displacement over
a period of motion lasting approximately 12 seconds. Details of
exactly how this displacement was measured are given below.

### 2011-04-13 paper_ecal/results.tex
More edits, 75% done :)

\subsection{Exploration of Parameterized Gait Space}

Before optimizing the chosen family of parameterized gaits
(\secref{motionModel}) with learning methods, we performed an
experiment to explore the five dimensions of the SineModel5 parameter
space. Specifically, we selected a random parameter vector that
resulted in some motion, but not an exceptional gait. We then varied
each of the five parameters individually and measured performance,
repeating each measurement twice to get a rough estimate of the
measurement noise at each point.  The results of this exploration,
shown in \figref{explore_dim_1}, reveal that some dimensions ($\amp$,
$\tau$, $m_F$) are fairly smooth and exhibit global structure across
the allowable parameter range, while others ($m_O$, $m_R$) exhibit
more complex behavior.  In addition, it gives a rough indication that
measurement noise is often significant and is more likely to be large
for gaits that move more.  Of course, this is only a slice in each
dimension through a single point, and slices through a different point
could reveal different behavior.  The common point at the
intersection of all slices is shown as a red triangle in
each plot of \figref{explore_dim_1}.

@@

\acmFiggggg{explore_dim_1}{explore_dim_2}{explore_dim_3}{explore_dim_4}{explore_dim_5}{1}{.53}{Fitness
  mean and standard deviation when each parameter dimension is varied
  independently.  The red triangle in each plot represents the same point in the 5-dimensional parameter space.}

>>>>

% \subsection{Exploration of Parameterized Gait Space}
% 
% Before optimizing the chosen family of parameterized gaits
% (\secref{motionModel}) with learning methods, we performed an
% experiment to explore the five dimensions of the SineModel5 parameter
% space. Specifically, we selected a random parameter vector that
% resulted in some motion, but not an exceptional gait. We then varied
% each of the five parameters individually and measured performance,
% repeating each measurement twice to get a rough estimate of the
% measurement noise at each point.  The results of this exploration,
% shown in \figref{explore_dim_1}, reveal that some dimensions ($\amp$,
% $\tau$, $m_F$) are fairly smooth and exhibit global structure across
% the allowable parameter range, while others ($m_O$, $m_R$) exhibit
% more complex behavior.  In addition, it gives a rough indication that
% measurement noise is often significant and is more likely to be large
% for gaits that move more.  Of course, this is only a slice in each
% dimension through a single point, and slices through a different point
% could reveal different behavior.  The common point at the
% intersection of all slices is shown as a red triangle in
% each plot of \figref{explore_dim_1}.

@@

% \acmFiggggg{explore_dim_1}{explore_dim_2}{explore_dim_3}{explore_dim_4}{explore_dim_5}{1}{.53}{Fitness
%   mean and standard deviation when each parameter dimension is varied
%   independently.  The red triangle in each plot represents the same point in the 5-dimensional parameter space.}
